---
title: "Australian Rental Market in 2026"
author: "Mili Shah"
date: "March 16, 2026"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Set up

```{r}
#| label: load-packages-data
#| message: false
#| warning: false

# load packages
library(tidyverse)
library(tidymodels)
library(knitr)  
library(readr)
library(patchwork)

# load data sets
rentalmarket <- read_csv("data/australian_rental_market_2026.csv")

# observe where na exists
sum(is.na(rentalmarket))
colSums(is.na(rentalmarket))

# remove duplicate rows
rentalmarket_unique <- unique(rentalmarket)
rentalmarket |> count(street_address) |> arrange(desc(n))

# see which row was duplicated
rentalmarket |> 
  filter(duplicated(rentalmarket))

# 2 things with longitude: 145.7229, latitude: -17.01989
```

```{r}
#| label: EDA

rentalmarket |>
  ggplot(aes(x=price_display)) +
  geom_density()

# rentalmarket |>
#   ggplot(aes(y=price_display, x = bathrooms)) +
#   geom_point() +
#   labs(title = "Price x Bathrooms")
# 
# rentalmarket |>
#   ggplot(aes(y=price_display, x = bedrooms)) +
#   geom_point() +
#   labs(title = "Price x Bedrooms")
# 
# rentalmarket |>
#   ggplot(aes(y=price_display, x = parking_spaces)) +
#   geom_point() +
#   labs(title = "Price x Parking Spaces")

rentalmarket |>
  ggplot(aes(y = price_display)) +
  geom_boxplot()

median(rentalmarket$price_display)
```

```{r}
library(ozmaps)
library(ggplot2)
library(sf)
library(corrplot)

sf_oz <- ozmap("states")

ggplot() +
  geom_sf(data = sf_oz, fill = "lightblue", color = "black") +
  geom_point(data = rentalmarket_unique,
             aes(x = longitude, y = latitude, color = price_display),
             size = 3) +
  scale_color_viridis_c(option = "plasma", name = "Price") +
  labs(title = "Price Mapping") +
  theme_void() +
  coord_sf()

num_rentalmarket <- rentalmarket |>
  select(price_display,
         latitude,
         longitude,
         postcode,
         bathrooms,
         bedrooms,
         parking_spaces)
cor_matrix <- cor(num_rentalmarket, method = "spearman")

corrplot(cor_matrix, method = "color", type = "upper", # Display upper triangle only
         order = "hclust", # Cluster variables by correlation similarity
         tl.col = "black", tl.srt = 45, # Text label color and rotation
         addCoef.col = "black", # Add coefficients to the plot
         sig.level = 0.05, insig = "blank") # Leave blank non-significant correlations

```

Analyze this public dataset containing information on a housing rental market. Conduct your analysis with an original script written by you using any appropriate tool, including but not limited to Python or R: [Australian Rental Market Data 2026](https://www.kaggle.com/datasets/kanchana1990/australian-rental-market-data-2026)

1.  Describe the quality of the data and clean it appropriately.

2.  Analyze an important aspect of the data using a supervised or unsupervised model and justify your modeling choices using statistical criteria.

3.  Construct a visualization conveying the implications or impact of your model.

4.  Describe potential next steps, considering both value and level of effort.

\pagebreak
